---
title: "lab10"
author: "Zijing"
format: pdf
---

```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

# Q1

```{r}
nrow(candy)
```
There are 85 candy types in this dataset.

# Q2

```{r}
sum(candy$fruity)
```
There are 38 fruity candy types in this dataset. 

# Q3

```{r}
candy["HersheyÕs Kisses", ]$winpercent
```
My favorite candy is HersheyÕs Kisses and its winpercent value is 55.37545.

# Q4

```{r}
candy["Kit Kat", ]$winpercent
```
The winpercent for Kit Kat is 76.7686.

# Q5

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
Tootsie Roll Snack Bars' winpercent is 49.6535.

```{r}
library("skimr")
skim(candy)
```

# Q6
winpercent is on a different scale compared to all the other variables, which are all from 0 to 1. Sugarprecent and pricepercent are also relatively different from others as they include values between 0 and 1 , while other variables are just exclusively 0 or 1.

# Q7
In candy$chocolate column, 0 means the candy is not chocolaty, 1 means the candy is chocolaty.

```{r}
library(ggplot2)
```

# Q8
```{r}
ggplot(candy, aes(winpercent))+
  geom_histogram()
```

# Q9
The distribution is not symmetrical. Is is positively skewed.

# Q10
The center of the distribution is below 50%.

# Q11
```{r}
choc <- candy$winpercent[as.logical(candy$chocolate)]
fruit <- candy$winpercent[as.logical(candy$fruity)]
mean(choc)>mean(fruit)
```
On average chocolate candy is higher ranked than fruit candy.

# Q12
```{r}
t.test(choc, fruit)
```
The p value is smaller than 0.001. The difference is statistically significant.

```{r}
library(dplyr)
```

# Q13
```{r}
candy %>% arrange(winpercent) %>% head(5)
```
Above are the 5 least liked candy type in this set.

# Q14
```{r}
candy %>% arrange(winpercent) %>% tail(5)
```
Above are the 5 favorite candy types in this set.

# Q15

```{r}
ggplot(candy, aes(winpercent, rownames(candy)))+
  geom_col()
```

# Q16
```{r}

ggplot(candy, aes(winpercent, reorder(rownames(candy),winpercent)))+
  geom_col()
```


```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

# Q17
The lowest ranked chocolate candy is sixlets.

# Q18
The highest ranked fruit candy is Starburst.

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

# Q19
```{r}
ord <- order(candy$winpercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )

```
ReeseÕs Miniatures is the highest ranked in terms of winpercent for the least money 

# Q20
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Above are the 5 most expensive candy type and Nik L Nip is the least popular one.

# Q21
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```
# Q22
Chocolate and fruity are the most anti-correlated, the second one is pluribus and bar.

# Q23
Chocolate and winpercent are the most positively correlated

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
library(plotly)
```
```{r}
#ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

# Q24. 
Fruity, pluribus are picked up by PC1 most strongly in the positive direction. This make sense because fruity candy usually come in a bag with multiple candies. Hard is also picked up in the positive direction, but not that strongly. 
