---
title: "lab17"
output: 
  pdf_document: 
    latex_engine: xelatex
author: "Zijing"
date: "2022-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(skimr)
```

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)

```
# Q1. What column details the total number of people fully vaccinated?
persons_fully_vaccinated

# Q2. What column details the Zip code tabulation area?
zip_code_tabulation_area

# Q3. What is the earliest date in this dataset?
2021-01-05

# Q4. What is the latest date in this dataset?
2022-11-22

```{r}
skimr::skim(vax)
```
# Q5. How many numeric columns are in this dataset? 
13

# Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
14921

# Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 
8.54

# Q8. [Optional]: Why might this data be missing?
Missing data from some zip code areas. 
```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
# Q9. How many days have passed since the last update of the dataset? 
4 days

# Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
99 unique dates
```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```
# Q11. How many distinct zip codes are listed for San Diego County? 
```{r}
length(unique(sd$zip_code_tabulation_area))
```
# Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
```{r}
zip <- which.max(sd$age12_plus_population)
sd$zip_code_tabulation_area[zip]
```
# Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-11-15”? 
```{r}
sd_221115 <- filter(sd, as_of_date == "2022-11-15")
mean(sd_221115$percent_of_population_fully_vaccinated, na.rm = TRUE)
```
Average percentage is 73.7%

# Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-11-15”?
```{r}
library(ggplot2)
```

```{r}
ggplot(sd_221115, aes(percent_of_population_fully_vaccinated))+
  geom_histogram()
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
# Q15
```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",title="Vaccination Rate for La Jolla CA 92037")
```

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-11-15")
```

```{r}
comp_mean <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
comp_mean
```
# Q16
```{r}
ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  geom_hline(yintercept = comp_mean, linetype = "dashed", color="red")+
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated",title="Vaccination Rate for La Jolla CA 92037")
```
# Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-11-15”?
```{r}
skimr::skim(vax.36)
```
Min 3.78501e-01, 1st Qu. 6.396185e-01, Median 7.15524e-01, Mean 7.172851e-01, 3rd Qu. 7.879820e-01, and Max 1.00000e+00

# Q18
```{r}
ggplot(vax.36, aes(percent_of_population_fully_vaccinated))+
  geom_histogram()+
  xlim(0,1)
```
# Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?
```{r}
vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2022-11-15") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
Both areas are below the average calculated above.

# Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(0.00,1.00) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = comp_mean, linetype="dashed")
```

