---
title: "lab6_hw"
output: pdf_document
author: "Zijing"
date: '2022-10-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bio3d)
```


## Q6

This function evaluates the b-factor similarity of different proteins. Enter the wanted chain and elety to be compared and the proteins being compared as a vector. The function would output a hierarchical cluster dendrogram based on protein b-factor similarity.

Inputs: 
chain: the chain identifier that user wants to compare within proteins 
elety: atom type to be compared within proteins
proteins: the list of proteins to be compared as a vector

Output: a hierarchical cluster dendrogram showing b-factor correlation among proteins.
```{r}
protein_comp <- function(chain,elety,proteins){
  labs <- c()
  for (i in 1:length(proteins)){
    protein <- read.pdb(proteins[i])
    protein.subset <- trim.pdb(protein, chain=chain,elety=elety)
    protein.b <- (protein.subset$atom$b)
    labs <- c(labs,proteins[i])
    if (i==1){
      protein.list <- protein.b
    }
    else{
      protein.list <- rbind(protein.list,protein.b)
    }
  }
  hc <- hclust(dist(protein.list))
  return(plot(hc,labels=labs))
}

```

```{r}
protein_comp("A","CA",c("4AKE","1AKE","1E4Y"))
```

