---
title: "lab12"
output: pdf_document
author: "Zijing"
date: "2022-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(BiocManager)
library(DESeq2)
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", row.names=1)
metadata <-  read.csv("airway_metadata.csv")
head(counts)
head(metadata)
```
# Q1
```{r}
nrow(counts)
```
There are 38694 genes

# Q2

```{r}
sum(metadata$dex == "control")
```
There are 4 control cell lines.

```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.counts <- counts[ ,control$id]
control.mean <- rowSums( control.counts )/4 
head(control.mean)
```

```{r}
library(dplyr)
control <- metadata %>% filter(dex=="control")
control.counts <- counts %>% select(control$id) 
control.mean <- rowSums(control.counts)/4
head(control.mean)
```

# Q3
Change the number 4 to sum(metadata$dex == "control")

# Q4
```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.mean <- rowSums( counts[ ,treated$id] )/4 
names(treated.mean) <- rownames(counts)
head(treated.mean)
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

# Q5a

```{r}
plot(meancounts$control.mean,meancounts$treated.mean)
```

```{r}
library(ggplot2)
```

# Q5b

```{r}
ggplot(meancounts,aes(control.mean,treated.mean))+
  geom_point()
```

# Q6

```{r}
plot(meancounts$control.mean,meancounts$treated.mean,log="xy")
```

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

# Q7
arr.ind is used so that array indices are returned. The first column of the output is the row indices, pointing to genes. The unique is used because sometimes a gene can have zero expression in both control.mean and treated.mean, leading to a repeat of its index in the first column of the output. Unique makes sure any repeat is eliminated, making sure we are not excluding wanted data points.

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

# Q8
```{r}
sum(up.ind)
```
250 up regulated genes at the greater than 2 fc level.

# Q9
```{r}
sum(down.ind)
```
367 down regulated genes at the greater than 2 fc level.

# Q10
No, we are not sure if the up/down regulations are statistically significant or not.

```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="SYMBOL", 
                     multiVals="first")
```

```{r}
head(res)
```

# Q11
```{r}
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="ENTREZID", 
                     multiVals="first")

res$uniport <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="UNIPROT", 
                     multiVals="first")

res$genename <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="GENENAME", 
                     multiVals="first")
head(res)
```

```{r}
ord <- order( res$padj )
head(res[ord,])
```

```{r}
write.csv(res[ord,], "deseq_results.csv")
```

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
```

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")

abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)
```

```{r}
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```

```{r}
library(EnhancedVolcano)
```

```{r}
x <- as.data.frame(res)

EnhancedVolcano(x,
    lab = x$symbol,
    x = 'log2FoldChange',
    y = 'pvalue')
```

```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)

head(kegg.sets.hs, 2)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
attributes(keggres)
```

```{r}
head(keggres$less, 3)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa05310")
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa05310", kegg.native=FALSE)
```

# Q12
```{r}
head(keggres$greater, 3)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa00500", kegg.native=FALSE)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa00330", kegg.native=FALSE)
```

